---
title: "Lab 04 - Visualizing spatial data"
author: "Allison"
date: "02052025"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
##install.packages("devtools")
##devtools::install_github("rstudio-education/dsbox")
library(tidyverse)
library(dsbox)
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

### Exercise 1

```{r descriptives Dennys dataset}
print(dennys)
```

There are `r nrow(dennys)` observations and `r ncol(dennys)` variables in the dataset. Each row in the dennys dataset represents a Denny's restaurant at different places. Each column represents one type of information about the location of the store, including the state, the city, the store specific location, ZIP code, the longitude, and latitude. 

### Exercise 2

```{r descriptives La Quintas dataset}
print(laquinta)
```

There are `r nrow(laquinta)` observations and `r ncol(laquinta)` variables in this dataset. Each row in the La Quinta's dataset represents a La Quinta hotel at different places. Each column represents one type of information about the location of the store, including the state, the city, the store specific location, ZIP code, the longitude, and latitude. 


### Exercise 3

On the website of Denny's it shows that the restaurant is opened in 15 countries besides USA, including: Canada, Puerto Rico, Guam, Mexico, Honduras, Costa Rica, Guatemala, El Salvador, The Philippines, Indonesia, New Zealand, The United Arab Emirates, The United Kingdom, Japan. For La Quinta's, the hotel is opened in 9 countries besides USA, including: Ecuador, Colombia, Chile, United Arab Emirates, Turkey, New Zealand, China, Mexico, Canada. 

### Exercise 4

if possible, we can create a new variable and filter the original state column into the new variable. The state abbreviations for the 50 states will be filtered as "US", while any other labels will be "others". In this way we can know how many of the stores are outside USA.    

### Exercise 5

```{r filter countries outside US}
dn <- dennys %>% 
  filter(!(state %in% states$abbreviation))
```

For Denny's, it appears that the dataset did not include any store that is outside of the US. 

### Exercise 6

```{r new variables for dennys}
dennys %>%
  mutate(country = "United States")

print(dn)
```

### Exercise 7

```{r laquinta countries outside the US}
lq <- laquinta %>% 
  filter(!(state %in% states$abbreviation))
print(lq)
```

According to the dataset, there are 14 stores outside of the US. According to the websites, the hotel is opened in 9 countries besides USA, including: Ecuador, Colombia, Chile, United Arab Emirates, Turkey, New Zealand, China, Mexico, Canada. 

### Exercise 8

```{r country variable}
laquinta <- laquinta %>%
  mutate(country = case_when(
    state %in% state.abb ~ "United States",
    state %in% c("ON", "BC") ~ "Canada",
    state == "ANT" ~ "Colombia",
    state %in% c("AG", "QR", "CH", "NL", "VE", "PU", "SL") ~ "Mexico",
    state == "FM" ~ "Honduras"
  ))

lqf <- laquinta %>%
  filter(country == "United States")
```

### Exercise 9

```{r how many stores}
dn %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation"))

lqf %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation"))
```
Based on the dataset, California has the most Dennys' stores (n = 403) and Delaware has the least store (n = 1). This is not surprising since Dennys started in California.
As for La Quinta hotel, Texas has the most La Quinta hotels (n = 237) and Maine has the least hotel (n = 1). The internet said that La Quinta started in Texas. 

### Exercise 10

```{r establishment}
dn <- dn %>%
  mutate(establishment = "Denny's")
lqf <- lqf %>%
  mutate(establishment = "La Quinta")
dn_lq <- bind_rows(dn, lqf)

ggplot(dn_lq, mapping = aes(
  x = longitude,
  y = latitude,
  color = establishment
)) +
  geom_point()
```


